---
title: "Pinelli_Gisler"
author: "Gianni Pinelli and Marius Gisler"
date: "2 10 2020"
output:
  pdf_document: default
  html_document: default
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

### Install Packages
```{r}
library(dplyr)
library(tinytex)
library(ggplot2)
library(ggrepel)
library(gridExtra)
library(gganimate)
library(gifski)
library(av)
```


```{r, include=FALSE}
getwd()
df <- read.csv('C:/Users/mariu/OneDrive - Hochschule Luzern/r_bootcamp/2020_Results.csv', header =TRUE, sep=";")
df <- data.frame(df)
```

### Data Cleaning
```{r}
names(df)[names(df) == "NSA.Code"] <- "Nation"
names(df)[names(df) == "ï..Rank" ] <- "Rank"
names(df)[names(df) == "Year.of.Birth" ] <- "Year_of_birth"
names(df)[names(df) == "Distance..m." ] <- "Distance_in_meters"
names(df)[names(df) == "Race.Points" ] <- "Race_points"
names(df)[names(df) == "Race.Day" ] <- "Race_day"
names(df)[names(df) == "FIS.Code" ] <- "FIS_code"
names(df)[names(df) == "Last.Name" ] <- "Last_name"
names(df)[names(df) == "First.Name" ] <- "First_name"
df$FIS_code <- as.factor(df$FIS_code)
df$Race_day <- as.factor(df$Race_day)
df$Race_day <- as.Date(df$Race_day, "%d.%m.%Y")
df$Race_year <- format(df$Race_day,'%Y')
df$Race_year <- as.numeric(df$Race_year)

```


### Add Additional Columns
```{r}
###Number of participations
df_part <- table(df$FIS_code)
df_part <- as.data.frame(df_part)
names(df_part)[names(df_part) == "Var1"] <- "FIS_code"
names(df_part)[names(df_part) == "Freq"] <- "Number_of_participations"
df <- left_join(df, df_part,by='FIS_code', keep=FALSE)

###Age of Rider at Race
df$year_of_race <- format(as.Date(df$Race_day, format="%d.%m.%Y"),"%Y")
df$year_of_race <- as.integer(df$year_of_race)
df$Age_of_rider <- df$year_of_race - df$Year_of_birth

###Number of wins
df_winner <- filter(df, df$Rank == 1)
df_winner_count <- table(df_winner$FIS_code)
df_winner_count <- as.data.frame(df_winner_count)
names(df_winner_count)[names(df_winner_count) == "Var1"] <- "FIS_code"
names(df_winner_count)[names(df_winner_count) == "Freq"] <- "Number_of_wins"
df <- left_join(df, df_winner_count,by='FIS_code', keep=FALSE)

###Number of wins/nation
df_winner_count_nation <- table(df_winner$Nation)
df_winner_count_nation <- as.data.frame(df_winner_count_nation)

names(df_winner_count_nation)[names(df_winner_count_nation) == "Var1"] <- "Nation"
names(df_winner_count_nation)[names(df_winner_count_nation) == "Freq"] <- "Number_of_wins_nation"
df <- left_join(df, df_winner_count_nation,by='Nation', keep=FALSE)
df$Number_of_wins_nation[is.na(df$Number_of_wins_nation)] <- 0

### Number of Participations / nation
nations <- table(df$Nation)
nations <- as.data.frame(nations)
names(nations)[names(nations) == "Var1"] <- "Nation"
names(nations)[names(nations) == "Freq"] <- "Number_of_participations_Nation"
df <- left_join(df, nations,by='Nation', keep=FALSE)

###Number of wins/ski
df_winner_ski <- table(df_winner$Ski)
df_winner_ski <- as.data.frame(df_winner_ski)
names(df_winner_ski)[names(df_winner_ski) == "Var1"] <- "Ski"
names(df_winner_ski)[names(df_winner_ski) == "Freq"] <- "Number_of_wins_ski"
df <- left_join(df, df_winner_ski,by='Ski', keep=FALSE)
df$Number_of_wins_ski[is.na(df$Number_of_wins_ski)] <- 0


### Number of Participations / ski
Ski <- table(df$Ski)
Ski <- as.data.frame(Ski)
names(Ski)[names(Ski) == "Var1"] <- "Ski"
names(Ski)[names(Ski) == "Freq"] <- "Number_of_participations_Ski"
df <- left_join(df, Ski,by='Ski', keep=FALSE)


###Number of podiums
df_podium <- filter(df, df$Rank <4)
df_podium <- table(df_podium$FIS_code)
df_podium <- as.data.frame(df_podium)
names(df_podium)[names(df_podium) == "Var1"] <- "FIS_code"
names(df_podium)[names(df_podium) == "Freq"] <- "Number_of_podiums"
df <- left_join(df, df_podium,by='FIS_code', keep=FALSE)

###Best Rank

df_rank =  df %>%
  select(FIS_code, Rank, Last_name)
df_rank <- df_rank[order(df_rank$Rank),]
df_rank <- df_rank[!duplicated(df_rank$FIS_code), ]
names(df_rank)[names(df_rank) == "Rank"] <- "Best_Rank"
df_rank$Last_name <- NULL
df <- left_join(df, df_rank,by='FIS_code', keep=FALSE)

head(df)
```



### Add additional Dataframes
```{r}
### Dataframe Podium
df_podium <- filter(df, df$Rank <4)

### Dataframe Winner

df_winner <- filter(df, df$Rank == 1)

### Dataframe Top 10
df_top_10 <- filter(df, df$Rank <11)

### Riders with more than 5 Participations
df_participations <- filter(df, df$Number_of_participations >6)
df_participations <- df_participations[order(df_participations$Rank),]
df_participations <- df_participations[!duplicated(df_participations$FIS_code), ]

```

### Data Visualisation: Nations most participated
```{r}
### Nations number of Participants
nations <- table(df$Nation)
nations <- as.data.frame(nations)
names(nations)[names(nations) == "Var1"] <- "Nation"
names(nations)[names(nations) == "Freq"] <- "Number_of_participations_Nation"

nations <- filter(nations, nations$Number_of_participations_Nation >9)
ggplot(data = nations, aes(x=reorder(Nation, -Number_of_participations_Nation), y=Number_of_participations_Nation)) +
  theme_minimal()+
  geom_bar(stat="identity", width=0.5, aes(fill = Nation))+
  ggtitle("Number of Participations per Nation")+
  xlab("Nation")+
  ylab("Number of Participations")+
  geom_text(aes(label=Number_of_participations_Nation), position=position_dodge(width=0.9), vjust=-0.25)+
  theme(axis.title.y=element_blank(),
        axis.text.y=element_blank(),
        axis.ticks.y=element_blank())
  
  
```


### Data Visualisation: Comparison participations, wins and wins per number of participations 
```{r}
df$Wins_per_Start_nation <- df$Number_of_wins_nation/df$Number_of_participations_Nation
nations <- filter(df, df$Number_of_participations_Nation >24)

plot_nations_1 <- ggplot(data = nations, aes(x=reorder(Nation, -Number_of_participations_Nation), y=Number_of_participations_Nation)) +
  theme_minimal()+
  geom_bar(stat="identity", width=0.5, position = "dodge", aes(fill = Nation),show.legend = FALSE)+
  ylab("Participations")+
  xlab("Nation")

plot_nations_2 <- ggplot(data = nations, aes(x=reorder(Nation, -Number_of_wins_nation), y=Number_of_wins_nation)) +
  theme_minimal()+
  geom_bar(stat="identity", width=0.5, position = "dodge", aes(fill = Nation),show.legend = FALSE)+
  ylab("Number of Wins")+
  xlab("Nation")

plot_nations_3 <- ggplot(data = nations, aes(x=reorder(Nation, -Wins_per_Start_nation), y=Wins_per_Start_nation)) +
  theme_minimal()+
  geom_bar(stat="identity", width=0.5, position = "dodge", aes(fill = Nation), show.legend = FALSE)+
  ylab("Wins per Start")+
  xlab("Nation")+ 
  scale_y_continuous(labels = scales::percent)

grid.arrange(plot_nations_1, plot_nations_2,plot_nations_3, ncol=1)


```

Austria has participated the most, however the Swiss won more races. Interestingly the Norwegians have a better rating in Wins divided by number of participations, this might indicate that the small norwegian team is strong or has had a lucky day by wining a race.




### Data Visualisation: Age and Rank over time

```{r}
ggplot(data = df, aes(x=Age_of_rider, y=Rank)) +
  geom_point()+
  coord_cartesian(ylim = c(1, 80))+
  scale_y_reverse()+
  geom_smooth(method = "lm")+
  theme_minimal()+
  xlab("Age")+
  ylab("Rank")

```


### Data Visualisation: Age Comparison

```{r}
age_average <- ggplot(data = df, aes(y=Age_of_rider))+
  geom_boxplot(color="black", fill="brown", alpha=0.2)+
  ylab("Age of all Athletes")+
  coord_cartesian(ylim = c(18, 40))+
  theme(axis.title.x=element_blank(),
        axis.text.x=element_blank(),
        axis.ticks.x=element_blank())
age_average

age_podium_average <- ggplot(data = df_podium, aes(y=Age_of_rider))+
  geom_boxplot(color="black", fill="gold", alpha=0.2)+
  ylab("Age of Athletes on Podium")+
  coord_cartesian(ylim = c(18, 40))+
  theme(axis.title.x=element_blank(),
        axis.text.x=element_blank(),
        axis.ticks.x=element_blank())
age_podium_average

age_winner_average <- ggplot(data = df_winner, aes(y=Age_of_rider, fill = Age_of_rider))+
  geom_boxplot(color="black", fill="green", alpha=0.2)+
  ylab("Age of winning Athletes")+
  coord_cartesian(ylim = c(18, 40))+
  theme(axis.title.x=element_blank(),
        axis.text.x=element_blank(),
        axis.ticks.x=element_blank())
age_winner_average

grid.arrange(age_average, age_podium_average, age_winner_average, ncol=3)

```
The age of the winning athletes and the athletes on the podium seem to be higher than the overall age of all athletes.




### Data Visualisation: Athletes with more than 6 starts, what is the highest rank reached?


```{r}
ggplot(df_participations, aes(y= Best_Rank, x = Number_of_participations)) + 
  geom_point(color = df_participations$Best_Rank, size = 3, show.legend = TRUE)+
  geom_label_repel(aes(label = Last_name),
                   box.padding   = 0.35, 
                   point.padding = 0.5,
                   segment.color = 'grey50',
                   size = 2) +
  theme(axis.text=element_text(size=9),
        axis.title=element_text(size=12,face="bold"))+
  coord_cartesian(xlim = c(5, 12))+
  scale_y_reverse()+
  ylab("Best rank")+
  xlab("Number of Starts")

```

### Data Analyisis: How do the nations perform on average?


```{r}
ggplot(data = nations, mapping = aes(x=reorder(Nation, Rank), y = Rank, fill=Nation))+
  geom_boxplot()+
  theme(axis.title.x=element_blank(),
        axis.text.x=element_blank(),
        axis.ticks.x=element_blank())

lm_nations <- lm(nations$Rank ~ nations$Nation)
summary(lm_nations)

```



### Data Analyisis: Are there correlations between Rank and Bib?


```{r}
cor(df$Rank, df$Bib)
ggplot(df, aes(x=Rank, y=Bib))+
  geom_point()+
  stat_smooth(method = 'lm', col = 'red')+
  ylab("Bib")+
  xlab("Rank")+
  ggtitle('Correlation between Rank and Bib')+
  theme_minimal()
lm.rank_1 <- lm(Rank ~ Bib, data = df)
summary(lm.rank_1)

```


### Data Analyisis: Are there correlations between Rank and Number of Participations?


```{r}
cor(df$Rank, df$Number_of_participations)
ggplot(df, aes(x=Rank, y=Number_of_participations))+
  geom_point()+
  stat_smooth(method = 'lm', col = 'red')+
  ylab("Number of Participations")+
  xlab("Rank")+
  ggtitle('Correlation between Rank and Bib')+
  theme_minimal()
lm.rank_2 <- lm(Rank ~ Number_of_participations, data = df)
summary(lm.rank_2)

```


### Comparison of models: Which one is better?

## A 10n fold cross validation is applied and the R-squards are compared.
```{r}
r.squared.simple <- c()
r.squared.few <- c()
set.seed(12)
for(i in 1:10){
  ## 1) prepare data
  smp_size <- floor(0.9*nrow(df))
  train <- sample(seq_len(nrow(df)), size = smp_size)
  df_train <- df[train,]
  df_test <- df[-train,]
  
  
  ## model 1
  ## 2) fit the model with "train" data
  lm.all <- lm(Rank ~ Bib + Number_of_participations, data = df)
  ##
  ## 3) make prediction on the test data
  predicted_lm_all <- predict(lm.all,
                              newdata = df_test)
  ##
  ## 4) compute R^2
  r.squared.simple[i] <- cor(predicted_lm_all,
                             df_test$Rank)^2
  
  
  
  ## model 2
  ## 2) fit the model with "train" data
  lm.few <- lm(Rank ~ Bib, data = df_train)
  ##
  ## 3) make prediction on the test data
  predicted_lm_few <- predict(lm.few,
                              newdata = df_test)
  ##
  
  ## 4) compute R^2
  r.squared.few[i] <- cor(predicted_lm_few,
                          df_test$Rank)^2
  }

mean_rsquared_lm_all <- mean(r.squared.simple)
mean_rsquared_lm_few <- mean(r.squared.few)
boxplot(r.squared.simple,r.squared.few)
  
  
print(mean_rsquared_lm_all)
print(mean_rsquared_lm_few)

```

The r-squared seem to do not differ significantly.



### Own Chapter

## Application of gganimate, gifski, and av.

```{r}

theme_set(theme_bw())

country <- c("SUI" = "red", "USA" = "blue", "ITA" = "green", "NOR" = "grey", "SLO" = "yellow", "LIE" = "orange", 
             "CAN" = "black", "AUT" = "pink", "SWE" = "beige", "FRA" = "brown", "GER" = "cyan" )
p <- ggplot(
  nations, 
  aes(x = Rank , y=Bib, size = Rank , colour = Nation)
) +
  geom_point(show.legend = TRUE, alpha = 0.7) +
  scale_color_viridis_d() +
  scale_size(range = c(2, 10)) +
  scale_x_log10() +
  labs(x = "Rank", y = "Bib") 
p


p + transition_time(Race_year) +
  labs(title = "Year: {frame_time}")

```